Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-06-14 Multiresolution_framework/ltx/manuscript.tex
edit to table3
minor edit to appendix
deleted old files
minor edits

B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig3.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (dotted lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

@@

	$\Delta$ & spatial discretisation step & 0.01 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 900 & n.a.\\

@@

	$\tau$ & membrane time constant & 0.01 & s$^{-1}$ \\
	$\varsigma$ & slope of sigmoidal activation function & 0.56 \citep{Wendling2005} & mV$^{-1}$.spike.s$^{-1}$\\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters (data generation) & $\left[\begin{array}{cc}

@@

	$n_y$ & number of sensors & 161 & n.a.\\ 
	$\Delta_y$ & distance between adjacent sensors & 0.05 & mm\\
	$\sigma_{m}$ & observation kernel full width at half maximum & 0.08 & mm \\
	$\Sigma_{\varepsilon}$ & observation noise covariance matrix & $0.1 \times I_{n_y}$ & mm$^2$ \\
	$\sigma_{\eta}$& disturbance covariance function full width at half maximum & 0.18 & mm\\
	$\sigma_{d}$ & disturbance variance & 0.53 & mV \\

@@

	$n_x$ & number of field basis functions&& n.a.\\
	$   $ &Experiment~I&131&\\
	$   $ &Experiment~II&17, 33, 65, 131, 263&\\ 

@@

	$n_{\theta}$&number of connectivity kernel basis functions for estimation&25&n.a.\\

@@

 		\includegraphics[scale=1]{./Graph/fig7.pdf}

@@

The EM algorithm was allowed to run until a maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Fig.~\ref{fig:MRA-Convergence}. 
\begin{figure}[t]
 	\centering
 		\includegraphics[scale=1]{./Graph/fig8.pdf}
 		\caption{{\bf Convergence of the EM based algorithm}. The mean  of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm over 200 realizations vs iterations of the EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.} 
\label{fig:MRA-Convergence}  
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states. 
The actual connectivity kernel and the decomposition is plotted in Fig.~\ref{fig:KernelEstimate}(A). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters needed to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. Figs.~\ref{fig:KernelEstimate}(B) and (C) illustrate the kernel scaling and wavelet functions, respectively. A total number of 25 basis functions were used to reconstruct the kernel, comprising 13 and 12 scaling and wavelet functions at $j=1$, respectively. 
\begin{figure}[!t]  
 \centering
\includegraphics{./Graph/fig9.pdf}
\caption{ {\bf Connectivity kernel estimate}. (\textbf{A}) The actual kernel and its components are shown by solid and dashed black lines, respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (\textbf{B},\textbf{C}) Kernel scaling and wavelets functions, respectively, at $j=1$ used in the estimation algorithm.}
\label{fig:KernelEstimate}
\end{figure}

@@

\includegraphics{./Graph/fig10.pdf}

@@

To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The RMSE of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.

@@

@@

		\includegraphics[scale=1]{./Graph/fig12.pdf}

@@

\begin{align}\label{eq:Xi0derivation0} 
 \mathbf M_{t+1}&=\mathbf E_{\theta'}\left[(\mathbf x_t-\mathbf{\hat x}_t^b)(\mathbf x_{t+1}-\mathbf{\hat x}_{t+1}^b)^\top\right] \nonumber \\
&=\mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\mathbf x_t\hat{\mathbf x}_{t+1}^{b\top}\right]-\mathbf E_{\theta'}\left[\hat{\mathbf x}_t^b\mathbf x_{t+1}^\top\right]+\mathbf E_{\theta'}\left[\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}\right]  \nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}+\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}\nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}.
\end{align}
Rearranging   Eq.~\eqref{eq:Xi0derivation0} gives
\begin{equation}\label{eq:Xi0derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]=\mathbf M_{t+1}+\mathbf {\hat x}_t^b\mathbf{\hat x}_{t+1}^{b\top}.
\end{equation}
Substituting Eq.~\eqref{eq:Xi0derivation1} into Eq.~\eqref{eq:appXi0}, noting the expectation commutes with the sum yields
\begin{equation}
 \boldsymbol\Xi_0=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t^b\mathbf{\hat x}_{t+1}^{b\top}\right)
\end{equation}   
An expression for $\boldsymbol\Xi_1$ in terms of the smoothed states can be also found in a similar way by calculating the covariance matrix, $\mathbf P_t^b$ and following the above steps.
%\begin{equation}\label{eq:Xi1derivation1}
% \mathbf E_{\theta'}\left[\mathbf x_{t}\mathbf x_{t}^\top\right]=\mathbf P_{t}+\mathbf {\hat x}_{t}\mathbf{\hat x}_{t}^\top,
%\end{equation}
%substituting \ref{eq:Xi1derivation1} into \ref{eq:appXi1} we get
%\begin{equation}
 %\boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf P_{t}+\mathbf{\hat x}_{t}\mathbf{\hat x}_{t}^\top\right).
%\end{equation}
\newpage
\bibliographystyle{./elsarticle/model2-names}  
\bibliography{MRAIDE}
\end{document}

>>>>

B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig3.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (dotted lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

\includegraphics[scale=1]{./Graph/fig5.pdf}

@@

	$\Delta$ & Spatial discretisation step & 0.01 & mm \\
	$T_s$ & Time step & $0.001$ & s\\
	$T$ & Number of time steps used in estimation & 900 & n.a.\\

@@

	$\tau$ & Membrane time constant & 0.01 & s$^{-1}$ \\
	$\varsigma$ & Slope of activation function & 0.56 \citep{Wendling2005} & mV$^{-1}$.spike.s$^{-1}$\\
	$\boldsymbol{\theta}$ & Vector of connectivity kernel parameters (data generation) & $\left[\begin{array}{cc}

@@

	$n_y$ & Number of sensors & 161 & n.a.\\ 
	$\Delta_y$ & Distance between adjacent sensors & 0.05 & mm\\
	$\sigma_{m}$ & Observation kernel full width at half maximum & 0.08 & mm \\
	$\Sigma_{\varepsilon}$ & Observation noise covariance matrix & $0.1 \times I_{n_y}$ & mm$^2$ \\
	$\sigma_{\eta}$& Disturbance covariance function full width at half maximum & 0.18 & mm\\
	$\sigma_{d}$ & Disturbance variance & 0.53 & mV \\

@@

	$n_x$ & Number of field basis functions&& n.a.\\
	$   $ &Experiment~I, $j=3$&131&\\
	$   $ &Experiment~II, $j=0,1,2,3,4$&17, 33, 65, 131, 263&\\ 

@@

	$n_{\theta}$&Number of connectivity kernel basis functions for estimation&25&n.a.\\

@@

@@

The EM algorithm was allowed to run until a maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Fig.~\ref{fig:MRA-Convergence}. 
\begin{figure}[t]
 	\centering
 		\includegraphics[scale=1]{./Graph/fig7.pdf}
 		\caption{{\bf Convergence of the EM based algorithm}. The mean  of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm over 200 realizations vs iterations of the EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.} 
\label{fig:MRA-Convergence}  
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states. 
The actual connectivity kernel and the decomposition is plotted in Fig.~\ref{fig:KernelEstimate}(A). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters needed to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. Figs.~\ref{fig:KernelEstimate}(B) and (C) illustrate the kernel scaling and wavelet functions, respectively. A total number of 25 basis functions were used to reconstruct the kernel, comprising 13 and 12 scaling and wavelet functions at $j=1$, respectively. 
\begin{figure}[!t]  
 \centering
\includegraphics{./Graph/fig8.pdf}
\caption{ {\bf Connectivity kernel estimate}. (\textbf{A}) The actual kernel and its components are shown by solid and dashed black lines, respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (\textbf{B},\textbf{C}) Kernel scaling and wavelets functions, respectively, at $j=1$ used in the estimation algorithm.}
\label{fig:KernelEstimate}
\end{figure}

@@

\includegraphics{./Graph/fig9.pdf}

@@

To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The mean over time of the root mean square error (MRMSE) over space of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.

@@

 		\includegraphics[scale=1]{./Graph/fig10.pdf}

@@

@@

\begin{align}\label{eq:Xi0derivation0} 
 \mathbf M_{t+1}&=\mathbf E_{\theta'}\left[(\mathbf x_t-\mathbf{\hat x}_t^b)(\mathbf x_{t+1}-\mathbf{\hat x}_{t+1}^b)^\top\right] \nonumber \\
&=\mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\mathbf x_t\hat{\mathbf x}_{t+1}^{b\top}\right]-\mathbf E_{\theta'}\left[\hat{\mathbf x}_t^b\mathbf x_{t+1}^\top\right]+\mathbf E_{\theta'}\left[\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}\right]  \nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}+\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}\nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t^b\hat{\mathbf x}_{t+1}^{b\top}.
\end{align}
Rearranging   Eq.~\eqref{eq:Xi0derivation0} gives
\begin{equation}\label{eq:Xi0derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]=\mathbf M_{t+1}+\mathbf {\hat x}_t^b\mathbf{\hat x}_{t+1}^{b\top}.
\end{equation}
Substituting Eq.~\eqref{eq:Xi0derivation1} into Eq.~\eqref{eq:appXi0}, noting the expectation commutes with the sum yields
\begin{equation}
 \boldsymbol\Xi_0=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t^b\mathbf{\hat x}_{t+1}^{b\top}\right).
\end{equation}   
An expression for $\boldsymbol\Xi_1$ in terms of the smoothed states can be also found in a similar way by calculating the covariance matrix, $\mathbf P_t^b$ and following the above steps.
%\begin{equation}\label{eq:Xi1derivation1}
% \mathbf E_{\theta'}\left[\mathbf x_{t}\mathbf x_{t}^\top\right]=\mathbf P_{t}+\mathbf {\hat x}_{t}\mathbf{\hat x}_{t}^\top,
%\end{equation}
%substituting \ref{eq:Xi1derivation1} into \ref{eq:appXi1} we get
%\begin{equation}
 %\boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf P_{t}+\mathbf{\hat x}_{t}\mathbf{\hat x}_{t}^\top\right).
%\end{equation}
\newpage
\bibliographystyle{./elsarticle/model2-names}  
\bibliography{MRAIDE}
\end{document}
